ANGSD-wrapper: utilities for analyzing next generation sequencing data

### 2016-01-25 manuscript.tex
for_submission

ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
ANGSD-wrapper provides a common configuration file, "common.conf," which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
Each analysis is run using a simple command:

@@

The visualization software included with ANGSD-wrapper is contained within it's own directory called "shinyGraphing."
This application must be started in R and can be accessed locally from a web browser. 
This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
Each tab in the GUI contains plots for different ANGSD methods.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5 megabase region of chromosome 10 in {\it Zea mays} spp. {\it mays} using ANGSD-wrapper. Blue boxes indicate genic regions provided by a GFF annotation. \label{fig:theta} }
\end{figure}

@@

As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}. The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}. In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.

We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp. \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples. 
The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X (Maize=0.0061, Teo=0.0082)\ad{wasn't sure if you wanted maize pi only or both so I included both}, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
Using the 2DSFS method, which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.   
There are no loci in this region that have previously been identified as showing evidence of selection during domestication, and these data also do not find any large consecutive regions of high \fst.

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
\caption{Admixture analysis for {\it Zea mays} spp. {\it mays}, {\it Zea mays} spp. {\it parviglumis}, and {\it Zea mays} spp. {mexicana} (\texit{mex}) with K=4 source populations. Individuals are shown in same order as Table \ref{tab:samples}. }
\label{fig:admixture}
\end{figure}

Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
Using the Admixture method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
\textit{Zea mays} ssp. {mexicana} clusters into its own group, along with a single accession of ssp. {parviglumis} collected from region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.  
A single  ssp. {parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.

>>>>

ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
ANGSD-wrapper provides a common configuration file, ``common.conf,'' which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
Each analysis is run using a simple command:

@@

The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
This application must be started in R and can be accessed locally from a web browser. 
This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
Each tab in the GUI contains plots for different ANGSD methods.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
\caption{Visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5Mb region of chromosome 10 in {\it Zea mays} ssp. {\it mays} using ANGSD-wrapper. Blue boxes indicate genic regions provided by a GFF annotation. \label{fig:theta} }
\end{figure}

@@

As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}. 
In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.

We first use the ``SFS'' method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp. \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples. 
The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.\jri{cite Eyre-Walker 1998 PNAS here instead}
Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with  mean levels of nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C). 
Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.   
There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis.

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
\caption{Admixture analysis for {\it Zea mays} ssp. {\it mays}, {\it Zea mays} ssp. {\it parviglumis}, and {\it Zea mays} ssp. {mexicana} (\textit{mex}) with K=4 source populations. 
Individuals are shown in the same order as Table \ref{tab:samples}. }
\label{fig:admixture}
\end{figure}

Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. \textit{mexicana} to assess evidence for admixture.  
Using the ``Admixture'' method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
We find no evidence of admixture between these lowland maize samples and  ssp. \textit{mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
\textit{Zea mays} ssp. \textit{mexicana} clusters into its own group, along with a single accession of ssp. \textit{parviglumis} collected from a region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.  
A single  ssp. \textit{parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
